---
title: "Communities and Crime"
subtitle: "Statistical Learning Final Exam Project"
chapter: "Data Cleaning"
author: "Caria Natascia, Cozzolino Claudia, Petrella Alfredo"
date: "June 20, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# set working directory
setwd("C:/Users/Natascia/Desktop/DataS_2/Stat_B/Project/code")
```

# Aim of the project

Predicting the rate of violent crime in a community can be particularly helpful in reducing the actual possibility of such crimes occurring.
Socio-economic, environmental and demographic characteristics can be important predictors of the level of violent crime in a population.
Therefore, determining which factors are the most influential can play a key role in understanding this complex phenomenon of crime.

This is what we are going to do with this project, by investigating the USA Communities and Crime Data Set, sourced from the [UCI Dataset Repository](https://archive.ics.uci.edu/ml/datasets/Communities+and+Crime+Unnormalized)[7].

# Dataset Description

The dataset contains a total number of 2215 instances, each one associated to a community in the US, and 147 attributes which result from the combination of socio-economic data from 1990 US Census, law enforcement data from the 1990 US LEMAS survey and crimedata from the 1995 FBI UCR.[^1]

[^1]: Attributes details in the appendix.

# Data Cleaning

```{r}
# import dataset (missing values stored as "?" - imported as NA)
crimedata <- read.csv("../data/crimedata.csv", na.strings="?")
dim(crimedata)
```

Attributes information:

* 4 non-predictive (communityname, countyCode, communityCode, fold)
* 125 predictive
* 18 potential response (murders, murdPerPop, rapes, rapesPerPop, robberies, robbbPerPop, assaults, assaultPerPop, burglaries, burglPerPop, larcenies, larcPerPop, autoTheft, autoTheftPerPop, arsons, arsonsPerPop, ViolentCrimesPerPop, nonViolPerPop)

From UCI description we know that the variables communityname and state are nominal while the remaining are all numeric.

```{r}
# check if variables communityname and state are stored as factors
is.factor(crimedata$communityname)
is.factor(crimedata$state)

# check number of numeric variables: 145 (= 147 - 2) expected
sum(sapply(crimedata, is.numeric))
```


## Missing values

```{r}
# inspect total NA
sum(is.na(crimedata))
mean(is.na(crimedata))
```

More than 10% of values is missing.

```{r, include=FALSE}
# install.packages("varhandle", dependencies = T)
library(varhandle)
```

```{r}
# dataframe with columns that present NA
nadf <- inspect.na(crimedata, hist=FALSE, summary=TRUE, byrow=FALSE, barplot=TRUE, 
                   na.value = NA)
```

There are 41 columns with missing values, many of them with more than 50% of the data missing.
Among them:

* 3 non-predictive (countyCode, communityCode, fold);
* many variables from US LEMAS dataset, including policing information;
* potential response (rapes, rapesPerPop, arsons, arsonsPerPop, ViolentCrimesPerPop, nonViolPerPop in particular)

From documentaion:

*The per capita violent crimes variable was calculated using population and the sum of crime variables considered violent crimes in the United States: murder, rape, robbery, and assault. There was apparently some controversy in some states concerning the counting of rapes. These resulted in missing values for rape, which resulted in missing values for per capita violent crime. Many of these omitted communities were from the midwestern USA (Minnesota, Illinois, and Michigan have many of these).*

*The per capita nonviolent crime variable was calculated using the sum of crime variables considered non-violent crimes in the United States: burglaries, larcenies, auto thefts and arsons. (There are many other types of crimes, these only include FBI 'Index Crimes').*

Looking to the dataset, other suspecious values equal to 0.00 have been found:

```{r}
zero_count <- function(x) sum(x==0, na.rm = TRUE)
zerodf <- lapply(crimedata, zero_count)[lapply(crimedata, zero_count)!=0]
names(zerodf)
```

Although it makes sense for some of these variables to take zero values, for others it seems unrealistic. However, we do not have any information about it in the documentation.

Therefore, columns with more than 50% NA are dropped and so are the variables fold (not useful), OwnOccQrange and RentQrange (obtained as the difference of other columns). We decide to keep the other columns with zero values due to their meaningfulness.

```{r}
# find columns with > 50% NA
coltodrop <- as.vector(nadf$column_name[nadf$ratio_of_NA>0.50])

# add other columns to remove
coltodrop <- c(coltodrop, "fold", "OwnOccQrange", "RentQrange")
coltodrop

# drop columns with > 50% NA and clearly redundant ones
cleandf <- crimedata[,!(names(crimedata) %in% coltodrop)]
```

### Possible ways to handle missing values

The remaining variables with missing values are the ones related to the crimes and another one, the variable OtherPerCap (per capita income for people with 'other' heritage), which has only one missing value.

```{r}
# remaining columns with NA
nadf <- inspect.na(cleandf, hist=FALSE, summary=TRUE, byrow=FALSE, barplot=FALSE,
                   na.value = NA)
nacolnames <- as.vector(nadf$column_name)
nacolnames
```

Possible ways to handle the remaining missing values:

1. drop all raws with at least one missing value;
2. substitute a missing value with the average computed over the state a community belongs;
3. leave NA value and consider it as another category.

```{r}
# rows with NA
narows <- inspect.na(cleandf, hist=FALSE, summary=TRUE, byrow=TRUE, barplot=FALSE,
                     na.value = NA)
dim(narows)[1]
```

Since there are a lot of rows that have at least one missing values, we decide to proceed as in point 2.

```{r}
# dataframe of the columns which still contain NA with the mean computed over the state
meandf <- aggregate(cleandf[,nacolnames], list(cleandf$state), 
                    function(x) mean(x, na.rm = TRUE))
sum(is.na(meandf))
```

The dataframe with the mean contains NA values, this means that there are States for which the value of a certain feature is zero for all communities that belong to them.
This happens for IL, MI (no data of rapes, rapesPerPop, ViolentCrimesPerPop), KS, VT (no data of arsons, arsonsPerPop, nonViolPerPop).
Moreover from documentation we know that also MN has a lot of missing values for rapes (59 out of 66):

```{r}
mnNA = sum(is.na(cleandf[cleandf$state=="MN","rapes"]))
mnTot = length(cleandf[cleandf$state=="MN","rapes"])
round(c(mnNA, mnTot, mnNA/mnTot), 2)
```

Therefore, data related to those states is removed.

```{r}
rowtodrop <- as.numeric(rownames(cleandf[cleandf$state %in% c("IL","MI","MN","KS","VT"),]))
cleandf <- cleandf[!(rownames(cleandf) %in% rowtodrop),]
```

For the remaining variables with missing values, we substitute NA with the mean computed over the state.

```{r}
# substitute a missing value with the average computed over the state
for(col in nacolnames) {
  match_table <- tapply(cleandf[[col]], cleandf$state, mean, na.rm=TRUE)
  NA_position <- which(is.na(cleandf[[col]]))
  cleandf[[col]][NA_position] <- match_table[cleandf$state[NA_position]]
}
```

Now cleandf does not contain NA.

```{r}
sum(is.na(cleandf))
dim(cleandf)
```

The resulting dataframe consists of 1996 instances and 120 attributes.

## Standardization

To avoid any bias due to the difference of the predictors content, the dataset must be somehow scaled.
Different methods have been tested because of the presence of several outliers, such as the classical mean-standard
deviation stardardization and the min-max normalization. In the latter case most of the information dropped lost for the previously
mentioned thickness of the tails of the columns densities, while we'll see that the former one, together with a logarithmic
transformation, gives far better results.

It is interesting to notice an additional detail: many columns of the dataset describe the same quantity for different categories.
An attempt has been to jointly standardize such groups of variables, considering common values for the mean and the standard deviation,
but the following results where basically equivalent to the first ones, so we decided not to overcomplicate the analysis.

```{r} 
standardization <- function(x) {
return ((x - mean(x)) / sd(x))
}

standf <- cleandf
standf[seq(3,dim(standf)[2])] <- lapply(standf[seq(3,dim(standf)[2])], standardization)
```

```{r, include=FALSE}
# remove all variables from environment excpet the cleaned and the original dataset
rm(list=setdiff(ls(), c("cleandf", "crimedata", "standf")))
```

```{r, include=FALSE}
# save cleandf and standf to CSV 
# write.csv(cleandf, "../data/crimedata-cleaned.csv", row.names=TRUE)
# write.csv(standf, "../data/crimedata-cleaned-stand.csv", row.names=TRUE)
```

\newpage

# Appendix

## Attributes details

125 predictive, 4 non-predictive, 18 potential responses

* communityname: Community name - not predictive - for information only (string)
* state: US state (by 2 letter postal abbreviation)(nominal)
* countyCode: numeric code for county - not predictive, and many missing values (numeric)
* communityCode: numeric code for community - not predictive and many missing values (numeric)
* fold: fold number for non-random 10 fold cross validation, potentially useful for debugging, paired tests - not predictive 
(numeric - integer)

#### Socio-economic data from 1990 US Census

Macro areas: race, age, income, employment, economic status, marital status, family structure, immigration data and home ownership

* population: population for community: (numeric - expected to be integer)
* householdsize: mean people per household (numeric - decimal)
* racepctblack: percentage of population that is african american (numeric - decimal)
* racePctWhite: percentage of population that is caucasian (numeric - decimal)
* racePctAsian: percentage of population that is of asian heritage (numeric - decimal)
* racePctHisp: percentage of population that is of hispanic heritage (numeric - decimal)
* agePct12t21: percentage of population that is 12-21 in age (numeric - decimal)
* agePct12t29: percentage of population that is 12-29 in age (numeric - decimal)
* agePct16t24: percentage of population that is 16-24 in age (numeric - decimal)
* agePct65up: percentage of population that is 65 and over in age (numeric - decimal)
* numbUrban: number of people living in areas classified as urban (numeric - expected to be integer)
* pctUrban: percentage of people living in areas classified as urban (numeric - decimal)
* medIncome: median household income (numeric - may be integer)
* pctWWage: percentage of households with wage or salary income in 1989 (numeric - decimal)
* pctWFarmSelf: percentage of households with farm or self employment income in 1989 (numeric - decimal)
* pctWInvInc: percentage of households with investment / rent income in 1989 (numeric - decimal)
* pctWSocSec: percentage of households with social security income in 1989 (numeric - decimal)
* pctWPubAsst: percentage of households with public assistance income in 1989 (numeric - decimal)
* pctWRetire: percentage of households with retirement income in 1989 (numeric - decimal)
* medFamInc: median family income (differs from household income for non-family households) (numeric - may be integer)
* perCapInc: per capita income (numeric - decimal)
* whitePerCap: per capita income for caucasians (numeric - decimal)
* blackPerCap: per capita income for african americans (numeric - decimal)
* indianPerCap: per capita income for native americans (numeric - decimal)
* AsianPerCap: per capita income for people with asian heritage (numeric - decimal)
* OtherPerCap: per capita income for people with 'other' heritage (numeric - decimal)
* HispPerCap: per capita income for people with hispanic heritage (numeric - decimal)
* NumUnderPov: number of people under the poverty level (numeric - expected to be integer)
* PctPopUnderPov: percentage of people under the poverty level (numeric - decimal)
* PctLess9thGrade: percentage of people 25 and over with less than a 9th grade education (numeric - decimal)
* PctNotHSGrad: percentage of people 25 and over that are not high school graduates (numeric - decimal)
* PctBSorMore: percentage of people 25 and over with a bachelors degree or higher education (numeric - decimal)
* PctUnemployed: percentage of people 16 and over, in the labor force, and unemployed (numeric - decimal)
* PctEmploy: percentage of people 16 and over who are employed (numeric - decimal)
* PctEmplManu: percentage of people 16 and over who are employed in manufacturing (numeric - decimal)
* PctEmplProfServ: percentage of people 16 and over who are employed in professional services (numeric - decimal)
* PctOccupManu: percentage of people 16 and over who are employed in manufacturing (numeric - decimal)
* PctOccupMgmtProf: percentage of people 16 and over who are employed in management or professional occupations (numeric - decimal)
* MalePctDivorce: percentage of males who are divorced (numeric - decimal)
* MalePctNevMarr: percentage of males who have never married (numeric - decimal)
* FemalePctDiv: percentage of females who are divorced (numeric - decimal)
* TotalPctDiv: percentage of population who are divorced (numeric - decimal)
* PersPerFam: mean number of people per family (numeric - decimal)
* PctFam2Par: percentage of families (with kids) that are headed by two parents (numeric - decimal)
* PctKids2Par: percentage of kids in family housing with two parents (numeric - decimal)
* PctYoungKids2Par: percent of kids 4 and under in two parent households (numeric - decimal)
* PctTeen2Par: percent of kids age 12-17 in two parent households (numeric - decimal)
* PctWorkMomYoungKids: percentage of moms of kids 6 and under in labor force (numeric - decimal)
* PctWorkMom: percentage of moms of kids under 18 in labor force (numeric - decimal)
* NumKidsBornNeverMar: number of kids born to never married (numeric - expected to be integer)
* PctKidsBornNeverMar: percentage of kids born to never married (numeric - decimal)
* NumImmig: total number of people known to be foreign born (numeric - expected to be integer)
* PctImmigRecent: percentage of _immigrants_ who immigated within last 3 years (numeric - decimal)
* PctImmigRec5: percentage of _immigrants_ who immigated within last 5 years (numeric - decimal)
* PctImmigRec8: percentage of _immigrants_ who immigated within last 8 years (numeric - decimal)
* PctImmigRec10: percentage of _immigrants_ who immigated within last 10 years (numeric - decimal)
* PctRecentImmig: percent of _population_ who have immigrated within the last 3 years (numeric - decimal)
* PctRecImmig5: percent of _population_ who have immigrated within the last 5 years (numeric - decimal)
* PctRecImmig8: percent of _population_ who have immigrated within the last 8 years (numeric - decimal)
* PctRecImmig10: percent of _population_ who have immigrated within the last 10 years (numeric - decimal)
* PctSpeakEnglOnly: percent of people who speak only English (numeric - decimal)
* PctNotSpeakEnglWell: percent of people who do not speak English well (numeric - decimal)
* PctLargHouseFam: percent of family households that are large (6 or more) (numeric - decimal)
* PctLargHouseOccup: percent of all occupied households that are large (6 or more people) (numeric - decimal)
* PersPerOccupHous: mean persons per household (numeric - decimal)
* PersPerOwnOccHous: mean persons per owner occupied household (numeric - decimal)
* PersPerRentOccHous: mean persons per rental household (numeric - decimal)
* PctPersOwnOccup: percent of people in owner occupied households (numeric - decimal)
* PctPersDenseHous: percent of persons in dense housing (more than 1 person per room) (numeric - decimal)
* PctHousLess3BR: percent of housing units with less than 3 bedrooms (numeric - decimal)
* MedNumBR: median number of bedrooms (numeric - decimal)
* HousVacant: number of vacant households (numeric - expected to be integer)
* PctHousOccup: percent of housing occupied (numeric - decimal)
* PctHousOwnOcc: percent of households owner occupied (numeric - decimal)
* PctVacantBoarded: percent of vacant housing that is boarded up (numeric - decimal)
* PctVacMore6Mos: percent of vacant housing that has been vacant more than 6 months (numeric - decimal)
* MedYrHousBuilt: median year housing units built (numeric - may be integer)
* PctHousNoPhone: percent of occupied housing units without phone (in 1990, this was rare!) (numeric - decimal)
* PctWOFullPlumb: percent of housing without complete plumbing facilities (numeric - decimal)
* OwnOccLowQuart: owner occupied housing - lower quartile value (numeric - decimal)
* OwnOccMedVal: owner occupied housing - median value (numeric - decimal)
* OwnOccHiQuart: owner occupied housing - upper quartile value (numeric - decimal)
* OwnOccQrange: owner occupied housing - difference between upper quartile and lower quartile values (numeric - decimal)
* RentLowQ: rental housing - lower quartile rent (numeric - decimal)
* RentMedian: rental housing - median rent (Census variable H32B from file STF1A) (numeric - decimal)
* RentHighQ: rental housing - upper quartile rent (numeric - decimal)
* RentQrange: rental housing - difference between upper quartile and lower quartile rent (numeric - decimal)
* MedRent: median gross rent (Census variable H43A from file STF3A - includes utilities) (numeric - decimal)
* MedRentPctHousInc: median gross rent as a percentage of household income (numeric - decimal)
* MedOwnCostPctInc: median owners cost as a percentage of household income - for owners with a mortgage (numeric - decimal)
* MedOwnCostPctIncNoMtg: median owners cost as a percentage of household income - for owners without a mortgage (numeric - decimal)
* NumInShelters: number of people in homeless shelters (numeric - expected to be integer)
* NumStreet: number of homeless people counted in the street (numeric - expected to be integer)
* PctForeignBorn: percent of people foreign born (numeric - decimal)
* PctBornSameState: percent of people born in the same state as currently living (numeric - decimal)
* PctSameHouse85: percent of people living in the same house as in 1985 (5 years before) (numeric - decimal)
* PctSameCity85: percent of people living in the same city as in 1985 (5 years before) (numeric - decimal)
* PctSameState85: percent of people living in the same state as in 1985 (5 years before) (numeric - decimal)

#### Law enforcement data from the 1990 US LEMAS survey
* LemasSwornFT: number of sworn full time police officers (numeric - expected to be integer)
* LemasSwFTPerPop: sworn full time police officers per 100K population (numeric - decimal)
* LemasSwFTFieldOps: number of sworn full time police officers in field operations (on the street as opposed to administrative etc) (numeric - expected to be integer)
* LemasSwFTFieldPerPop: sworn full time police officers in field operations (on the street as opposed to administrative etc) per 100K 
population (numeric - decimal)
* LemasTotalReq: total requests for police (numeric - expected to be integer)
* LemasTotReqPerPop: total requests for police per 100K popuation (numeric - decimal)
* PolicReqPerOffic: total requests for police per police officer (numeric - decimal)
* PolicPerPop: police officers per 100K population (numeric - decimal)
* RacialMatchCommPol: a measure of the racial match between the community and the police force. High values indicate proportions in 
community and police force are similar (numeric - decimal)
* PctPolicWhite: percent of police that are caucasian (numeric - decimal)
* PctPolicBlack: percent of police that are african american (numeric - decimal)
* PctPolicHisp: percent of police that are hispanic (numeric - decimal)
* PctPolicAsian: percent of police that are asian (numeric - decimal)
* PctPolicMinor: percent of police that are minority of any kind (numeric - decimal)
* OfficAssgnDrugUnits: number of officers assigned to special drug units (numeric - expected to be integer)
* NumKindsDrugsSeiz: number of different kinds of drugs seized (numeric - expected to be integer)
* PolicAveOTWorked: police average overtime worked (numeric - decimal)
* LandArea: land area in square miles (numeric - decimal)
* PopDens: population density in persons per square mile (numeric - decimal)
* PctUsePubTrans: percent of people using public transit for commuting (numeric - decimal)
* PolicCars: number of police cars (numeric - expected to be integer)
* PolicOperBudg: police operating budget (numeric - may be integer)
* LemasPctPolicOnPatr: percent of sworn full time police officers on patrol (numeric - decimal)
* LemasGangUnitDeploy: gang unit deployed (numeric - integer - but really nominal - 0 means NO, 10 means YES, 5 means Part Time)
* LemasPctOfficDrugUn: percent of officers assigned to drug units (numeric - decimal)
* PolicBudgPerPop: police operating budget per population (numeric - decimal)

#### Crimedata from the 1995 FBI UCR
* murders: number of murders in 1995 (numeric - expected to be integer) potential GOAL attribute (to be predicted)
* murdPerPop: number of murders per 100K population (numeric - decimal) potential GOAL attribute (to be predicted)
* rapes: number of rapes in 1995 (numeric - expected to be integer) potential GOAL attribute (to be predicted)
* rapesPerPop: number of rapes per 100K population (numeric - decimal) potential GOAL attribute (to be predicted)
* robberies: number of robberies in 1995 (numeric - expected to be integer) potential GOAL attribute (to be predicted)
* robbbPerPop: number of robberies per 100K population (numeric - decimal) potential GOAL attribute (to be predicted)
* assaults: number of assaults in 1995 (numeric - expected to be integer) potential GOAL attribute (to be predicted)
* assaultPerPop: number of assaults per 100K population (numeric - decimal) potential GOAL attribute (to be predicted)
* burglaries: number of burglaries in 1995 (numeric - expected to be integer) potential GOAL attribute (to be predicted)
* burglPerPop: number of burglaries per 100K population (numeric - decimal) potential GOAL attribute (to be predicted)
* larcenies: number of larcenies in 1995 (numeric - expected to be integer) potential GOAL attribute (to be predicted)
* larcPerPop: number of larcenies per 100K population (numeric - decimal) potential GOAL attribute (to be predicted)
* autoTheft: number of auto thefts in 1995 (numeric - expected to be integer) potential GOAL attribute (to be predicted)
* autoTheftPerPop: number of auto thefts per 100K population (numeric - decimal) potential GOAL attribute (to be predicted)
* arsons: number of arsons in 1995 (numeric - expected to be integer) potential GOAL attribute (to be predicted)
* arsonsPerPop: number of arsons per 100K population (numeric - decimal) potential GOAL attribute (to be predicted)
* ViolentCrimesPerPop: total number of violent crimes per 100K popuation (numeric - decimal) GOAL attribute (to be predicted)
* nonViolPerPop: total number of non-violent crimes per 100K popuation (numeric - decimal) potential GOAL attribute (to be predicted)

\newpage

# References

[1] "An Introduction to Statistical Learning", G. James, D. Witten, T. Hastie and R. Tibshirani, Springer, 2013.

[2] "DEA History Book, 1876–1990" (drug usage & enforcement), US Department of Justice, 1991, USDoJ.gov, webpage: DoJ-DEA-History-1985-1990.

[3] "Guns and Violence: The Enduring Impact of Crack Cocaine Markets on Young Black Males", W.N. Evans, G. Garthwaite, T. Moore, 2018.

[4] "Measuring Crack Cocaine and Its Impact", Fryer, Roland. Harvard University Society of Fellows: 3, 66. Retrieved January 4, 2016.

[5] "The New Jim Crow: Mass Incarceration in the Age of Colorblindness", M. Alexander.

[6] http://www.disastercenter.com/crime/uscrime.htm

[7] https://archive.ics.uci.edu/ml/datasets/Communities+and+Crime+Unnormalized

[8] https://online.stat.psu.edu/stat462/node/171/.

[9] https://peltarion.com/knowledge-center/documentation/modeling-view/build-an-ai-model/loss-functions/mean-squared-logarithmic-error-(msle)

[10] https://towardsdatascience.com/the-mathematics-behind-principal-component-analysis-fff2d7f4b643

## Citation Request

U. S. Department of Commerce, Bureau of the Census, Census Of Population And Housing 1990 United States: 
Summary Tape File 1a & 3a (Computer Files),

U.S. Department Of Commerce, Bureau Of The Census Producer, Washington, DC and Inter-university Consortium for Political and Social 
Research Ann Arbor, Michigan. (1992)

U.S. Department of Justice, Bureau of Justice Statistics, Law Enforcement Management And Administrative Statistics (Computer File) U.S. 
Department Of Commerce, Bureau Of The Census Producer, Washington, DC and Inter-university Consortium for Political and Social Research 
Ann Arbor, Michigan. (1992)

U.S. Department of Justice, Federal Bureau of Investigation, Crime in the United States (Computer File) (1995)
